Formalizing Stack Safety as a Security Property

### 2021-07-07 submission3/main.tex
Move section 4 before section 2

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
%\apt{maybe: ``Stack safety is a testable security property''?}\bcp{That
%  would sacrifice a lot of oomph for a little bit of precision, IMO.  Prefer
%  to leave it out of the title but hit it strongly in the abstract.}%
What exactly does ``stack safety'' mean? The phrase is associated with a
variety of compiler,
run-time, and hardware mechanisms for protecting stack
memory.  But these mechanisms typically lack precise specifications,
relying instead on informal descriptions and examples of bad
behaviors that they prevent.

We propose a formal characterization
of stack safety, formulated with concepts from language-based security: a
combination of an integrity property (``the private
state in each caller's stack frame is held invariant by the callee''),
a confidentiality property (``the callee's behavior is insensitive to the
caller's private state''), and a well-bracketedness property (``each
callee returns control to its immediate caller'').
% These properties are sensible even in the presence of
% abnormal control flow.\apt{Is this important enough to mention here?}
% BCP: I think not in the abstract, but maybe in the intro.
% We further apply\apt{what does this mean? we didn't
%  invent it} a straightforward notion of
% ``well-bracketed control flow'' that restricts control flow to the
% typical pattern of matched calls and returns.
%
We use these properties to validate the stack-safety ``micro-policies''
proposed by~\citet{DBLP:conf/sp/RoesslerD18}.  Specifically, we check (with
property-based random testing) that Roessler and Dehon's ``eager''
micro-policy, which catches violations as early as possible, enforces a
simple ``stepwise'' variant of our properties and correctly detects several
broken variants, and that (a repaired version of) their more performant
``lazy'' micro-policy corresponds to a slightly weaker and more extensional
``observational'' variant of our properties.
%
% \BCP{I still don't think the next sentence really nails it.  And (sigh) I am
%   still not really clear yet on whether our definition applies / might apply
%   to other enforcement mechanisms or not.  Saying that our attacker model is
%   stronger does not help---that should make it easier for us to describe
%   their protections, not harder.  So I don't understand what we're saying
%   here.}%
% %
% While we do not rule out other enforcement mechanisms, our
% definitions are strongly informed by micropolicy systems, especially
% our attacker model, which is stronger than other enforcement mechanisms
% can enable.

@@

The call stack is a perennial target for low-level attacks, leading to
consequences ranging from leakage or corruption of private stack data to
control-flow hijacking. To prevent or detect such attacks, a profusion of
software and hardware protections have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTKJFP,Georges+21},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
  nick; there could be more}\bcp{Yes, going back to MIT days---we should
  include several more of these, if only to give readers the impression that
this is a well-studied mechanism (so formalizing its protections is
useful).}
\fi
%
The protections offered by such mechanisms are commonly described in terms
of concrete examples of attacks that they can prevent---corruption
of return addresses, buffer overflows, use of uninitialized variables,
etc.---leaving a more abstract characterization to the reader's intuition.
But these mechanisms can be quite intricate,
and they often require trading off protection and efficiency.
Thus, it would be useful to have a precise, generic, and formal
specification for stack
safety, both as a basis for comparing the security claims of different
enforcement techniques and for validating more rigorously that these claims
are met by particular implementations.

@@

For ease of exposition, our development initially assumes a single simple
stack with no sharing between callers and callees: all parameters and return
values are passed in registers.  Later, we show how to refine the properties
to allow passing of scalar stack data, granting a callee access to the
caller's data. Throughout, we note how we expect elements of the formalism
to interact with enforcement mechanisms beyond micro-policies.

@@

%% %\bcp{I worry that we are being a little
%% %  dishonest here: the stated advantage of the lazy policy is that it is more
%% %efficient, but our variant is {\em not} efficient!}
In summary, we offer the following contributions:

@@

We begin with a description of our setting (\cref{sec:setup}) and threat
model (\cref{sec:threat}), and an informal overview of our key ideas via
examples (\cref{sec:running-example}). \Cref{sec:relwork,sec:future} discuss related and future work.

\section{Setting}
\label{sec:setup}

@@

This section introduces the intuition behind the properties through a series
of examples. We show our examples both in C syntax and in corresponding assembly code for a
simple two-address machine. The address of each instruction is shown to its left.
The assembly code reflects a simple compilation model in which
local variables are stored in the stack frame.
For the moment, arguments and return values are passed in registers.
The machine is word-addressed. The stack grows upward in
memory and {\tt \%sp} points to the first unused word above the top of the stack.
The {\tt call} and {\tt ret} instructions push and pop the return address from the stack,
implicitly adjusting the stack pointer.
At the right of each example, we show the layout of stack memory just before
the specified instruction is executed; each box represents a memory word, and
addresses increase to the right.

@@

>>>>

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
%\apt{maybe: ``Stack safety is a testable security property''?}\bcp{That
%  would sacrifice a lot of oomph for a little bit of precision, IMO.  Prefer
%  to leave it out of the title but hit it strongly in the abstract.}%
What does ``stack safety'' mean, exactly? The phrase is associated with a
variety of compiler,
run-time, and hardware mechanisms for protecting stack
memory, but these mechanisms typically lack precise specifications,
relying instead on informal descriptions and examples of bad
behaviors that they prevent.

We propose a formal characterization
of stack safety, formulated with concepts from language-based security: a
combination of an integrity property (``the private
state in each caller's stack frame is held invariant by the callee''),
a confidentiality property (``the callee's behavior is insensitive to the
caller's private state''), and a well-bracketedness property (``each
callee returns control to its immediate caller'').
% These properties are sensible even in the presence of
% abnormal control flow.\apt{Is this important enough to mention here?}
% BCP: I think not in the abstract, but maybe in the intro.
% We further apply\apt{what does this mean? we didn't
%  invent it} a straightforward notion of
% ``well-bracketed control flow'' that restricts control flow to the
% typical pattern of matched calls and returns.

We use these properties to validate the stack-safety {\em micro-policies}
proposed by~\citet{DBLP:conf/sp/RoesslerD18}.  Specifically, we check (with
property-based random testing) that Roessler and Dehon's ``eager''
micro-policy, which catches violations as early as possible, enforces a
simple ``stepwise'' variant of our properties and correctly detects several
broken variants, and that (a repaired version of) their more performant
``lazy'' micro-policy corresponds to a slightly weaker and more extensional
observational variant of our properties.
%
% \BCP{I still don't think the next sentence really nails it.  And (sigh) I am
%   still not really clear yet on whether our definition applies / might apply
%   to other enforcement mechanisms or not.  Saying that our attacker model is
%   stronger does not help---that should make it easier for us to describe
%   their protections, not harder.  So I don't understand what we're saying
%   here.}%
% %
% While we do not rule out other enforcement mechanisms, our
% definitions are strongly informed by micropolicy systems, especially
% our attacker model, which is stronger than other enforcement mechanisms
% can enable.

@@

The call stack is a perennial target for low-level attacks, leading to a
range of dire consequences, from leakage or corruption of private stack data
to control-flow hijacking. To prevent or detect such attacks, a profusion of
software and hardware protections have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTKJFP,Georges+21},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
  nick; there could be more}\bcp{Yes, going back to MIT days---we should
  include several more of these, if only to give readers the impression that
this is a well-studied mechanism (so formalizing its protections is
useful).}
\fi

The protections offered by such mechanisms are commonly described in terms
of concrete examples of attacks that they can prevent---corruption
of return addresses, buffer overflows, use of uninitialized variables,
etc.---leaving a more abstract characterization to the reader's intuition.
But these mechanisms can be quite intricate,
and they often involve trading off protection and performance.\bcp{I'm a
  little uncomfortable about that point: to the extent that applying these
  mechanisms involves trading off performance and protection}
Thus, it would be useful to have a precise, generic, and formal
specification for stack
safety, as a basis both for comparing the security claims of different
enforcement techniques and for validating more rigorously that these claims
are met by particular implementations.

@@

For ease of exposition, our development initially assumes a single simple
stack with no sharing between callers and callees: all parameters and return
values are passed in registers.  Later, we show how to refine the properties
to allow passing of scalar stack data, granting a callee access to the
caller's data. Throughout, we note how we expect elements of the formalism
to interact with\bcp{``interact with''??} enforcement mechanisms beyond micro-policies.

@@

%% %\bcp{I worry that we are being a little
%% %  dishonest here: the stated advantage of the lazy policy is that it is more
%% %efficient, but our variant is {\em not} efficient!}
In the rest of the paper, we begin with an informal overview of our key ideas via
examples (\cref{sec:running-example}), followed by definitions of some key
concepts (\cref{sec:setup}) and a discussion of the threat
model we consider (\cref{sec:threat}). We then offer the following contributions:

@@

\Cref{sec:relwork,sec:future} discuss related and future work.

@@

This section introduces the intuition behind our formal stack safety properties through a series
of examples, each presented both in C syntax and in corresponding assembly code for a
simple two-address machine (see Figure~\ref{fig:int1}). The address of each
instruction is shown to its left.
The assembly code reflects a simple compilation model in which
local variables are stored in the stack frame.
For the moment, arguments and return values are passed in registers.
The machine is word-addressed. The stack grows upward in
memory and {\tt \%sp} points to the first unused word above the top of the stack.
The {\tt call} and {\tt ret} instructions push and pop the return address from the stack,
implicitly adjusting the stack pointer.
At the right of each example, we show the layout of stack memory just before
the specified instruction is executed; each box represents a memory word, and
addresses increase to the right.

@@

\section{Key Concepts}
\label{sec:setup}
